---
title: "pokemon"
author: "Camille Benoit"
date: "29/12/2020"
output: html_document
---
```{r}
.libPaths("C:/Program Files (x86)/R/win-library/4.0")
library(dplyr)
library(ggplot2)
```

## Loading and display

```{r}
data = read.csv("./project/Pokemon.csv", header = TRUE, sep = ",")
```

```{r}
head(data)
```

```{r}
summary(data)
```

## Some manipulations

```{r}
data$Number <- data$X
data <- data[,-1] %>% 
  select(Number, Name:Legendary)
data[["Type.2"]][data[["Type.2"]] == ""] <- "None"
data[["Legendary"]] <- data[["Legendary"]] %>% 
                          as.logical()
data$Generation <- as.factor(data$Generation)
```

We want to work with gen1 to gen4 without legendary pokemon.

```{r}
gen1 <- data %>% 
  filter(!grepl("Mega", Name, fixed = TRUE) & Generation == 1 & Legendary == FALSE)

gen2 <- data %>% 
  filter(!grepl("Mega", Name, fixed = TRUE) & Generation == 2 & Legendary == FALSE)

gen3 <- data %>% 
  filter(!grepl("Mega", Name, fixed = TRUE) & Generation == 3 & Legendary == FALSE)

gen1to3<- data %>% 
  filter(!grepl("Mega", Name, fixed = TRUE) & as.numeric(Generation) < 4 & Legendary == FALSE)
```

## Some ideas 

We want to compare some statistics between different generations / types.

```{r}
gen1 %>% ggplot(aes(x= Attack, y= Defense, color = Type.1)) + geom_point() + facet_wrap(Type.1 ~ .)
```

```{r}
gen2 %>% ggplot(aes(x= Attack, y= Defense, color = Type.1)) + geom_point() + facet_wrap(Type.1 ~ .)
```

```{r}
gen3 %>% ggplot(aes(x= Attack, y= Defense, color = Type.1)) + geom_point() + facet_wrap(Type.1 ~ .)
```

example : we want to compare water type

```{r}
gen1to3 %>% filter(Type.1 == "Water") %>% 
  ggplot(aes(x= Attack, y= Defense, color = Generation)) + geom_point()
```

or

```{r}
gen1to3 %>% filter(Type.1 == "Fire") %>% 
  ggplot(aes(x= Attack, y= Defense, color = Generation)) + geom_point() +
  facet_wrap(. ~ Generation)
```

Distribution of Attack by types over generation.
```{r}
gen1to3 %>% 
  ggplot(aes(x= HP, fill = Generation, color = Generation)) + 
  geom_density(alpha = 0.3) + guides(fill = F)
```

```{r}
gen1to3 %>% 
  ggplot(aes(x= Attack, fill = Generation, color = Generation)) + 
  geom_density(alpha = 0.3) + guides(fill = F)
```

```{r}
gen1to3 %>% 
  ggplot(aes(x= Defense, fill = Generation, color = Generation)) + 
  geom_density(alpha = 0.3) + guides(fill = F)
```

Means of HP, attack, defense, etc... by types over generations. Does it changes a lot ? Is there some types which are better ?

```{r}
gen1 %>% 
  select(Type.1, HP:Speed) %>% 
    group_by(Type.1) %>% 
      summarize_all(funs(if(is.numeric(.)) mean(.) else table(.)))
```

```{r}
gen2 %>% 
  select(Type.1, HP:Speed) %>% 
    group_by(Type.1) %>% 
      summarize_all(funs(if(is.numeric(.)) mean(.) else table(.)))
```

```{r}
gen3 %>% 
  select(Type.1, HP:Speed) %>% 
    group_by(Type.1) %>% 
      summarize_all(funs(if(is.numeric(.)) mean(.) else table(.)))
```

Number of pokemon by types for gen1.

```{r}
gen1 %>% ggplot(aes(x=Type.1, fill = Type.1)) + geom_bar()
```

Evolution of number of water types through generations.

```{r}
data %>% 
  filter(Type.1 == "Water") %>% 
  select(Type.1, Generation) %>% 
  ggplot(aes(x=Generation, fill = Generation)) + geom_bar()

```

Mean of attack for water types pokemon through generations.

```{r}
data %>% 
  filter(Type.1 == "Water") %>%  
  select(Type.1, Generation, Attack) %>%  
  group_by(Generation) %>%  
  summarise(meanAttack = mean(Attack), .groups = "drop") %>% 
  ggplot(aes(x=Generation, y= meanAttack, color = Generation)) + geom_bar(stat =  "identity")
```