---
title: "Tasks"
author: "Camille Benoit"
date: "03/12/2020"
output:
  html_document: default
  pdf_document: default
---

```{r}
.libPaths("C:/Program Files (x86)/R/win-library/4.0")
```

# Task number 1
### Function + for loop

Instructions are to write a function that :

- Takes data frame, row selection and column selection as arguments

- Subsets data according to the given selections. Possible selections are: by index/indexes, by name/names, by logical vector

- For each column of the subseted data performs some computations: for numeric data – any of sum(), mean(), sd(), etc and for non-numeric data (characters and logical vectors, factors) – frequency table

- Returns an output as a data frame and calculation results as a list (subset + summary)

Here is what I wrote for this function.

```{r}
my_function <- function(data, rows, columns){
  subset <- data[rows, columns, drop = FALSE]
  res <- list(subset)
  for (col in colnames(subset)){
    if (is.numeric(subset[[col]])){
      meancol <- list(mean(subset[[col]]))
      res <- c(res, meancol)
    } else {
      tablecol <- list(table(subset[[col]]))
      res <- c(res, tablecol)
    }
  }
  names(res) <- c("Subset", colnames(subset))
  res
}
```

I used iris dataframe to do some tests.

```{r}
head(iris)
```

Here are my results :

```{r}
my_function(iris, 1:50 , c("Petal.Length","Sepal.Width", "Species"))
```

```{r}
my_function(iris, c(1:25,55:65,110:122) , c(4,2,5))
```

```{r}
my_function(iris, 45:102 , 3:5)
```

# Task number 2
### Function + looping function

Instructions for this task are to rewrite my previous function using *apply(), to add formal argument to split the data and to add calculation function as a formal argument to the function.

Here is what I wrote.

```{r}
my_new_function <- function(data, rows, columns, f, splt){
  subset <- data[rows, columns, drop = FALSE]
  spltcol <- data[rows, splt]
  concat_subset <- cbind(subset, spltcol)
  res <- list(subset)
  for (col in colnames(subset)){
    if (is.numeric(subset[[col]])){
      newres <- list(tapply(X = concat_subset[[col]], INDEX = spltcol, FUN = f))
    }else{
      newres <- list(table(subset[[col]]))
    }
    res <- c(res, newres)
  }
  names(res) <- c("Subset" ,colnames(subset))
  res
}
```

And some examples :

```{r}
my_new_function(iris, 1:150, 2:4, sum, "Species")
```

```{r}
my_new_function(iris, 1:75, 1:3, mean, 5)
```

```{r}
my_new_function(iris, 6:23, c(1,2), max, 5)
```

# Task 3
### Weather's data tidying and cleaning 

What I have to do for this task is to explore, tidy and prepare the dataset weather.rds

#### Raw data

Here is the raw data, with no modification (first 15 rows).

```{r}
weather <- readRDS("weather.rds")
head(weather, n=15)
```

#### Instruction list

I see that we have two issues with our raw data.

1/ Our variables are in a column named "measure" and Column headers are values (X1 -> X31) instead of variable names

2/ The first column with only index is useless.

To deal with that I will use tidyverse library.

```{r}
library(tidyverse)
```

As we are working with weather and already have "Month" and "Year" column, I will assume that X1...X31 correspond to days. After exploring data, I saw that 31st days corresponding to months with only 30 days contain NA values.

First, I remove the first column which is useless.

```{r}
weather2 <- weather[, -1]
head(weather2, n=10)
```

Then I gather all measurements by days and I will delete NA values (na.rm = TRUE).

```{r}
weather_gather <- gather(weather2, day, measurement_value, X1:X31, na.rm = TRUE)
head(weather_gather, n = 10)
tail(weather_gather, n= 10)
```

I spread the column measure to have a column for each measure.

```{r}
tidy_weather <- spread(weather_gather, measure, measurement_value)
head(tidy_weather, n = 10)
```

Finally I have in my dataset :

- Each object / observation is presented in a different row

- Each measured variable is presented in a different column

- Each cell – an intersection of given row and column – presents a value of the given variable for the given observation

Here is the **final instruction list** :

```{R}
weather2 <- weather[, -1]
weather_gather <- gather(weather2, day, measurement_value, X1:X31, na.rm = TRUE)
tidy_weather <- spread(weather_gather, measure, measurement_value)
```

#### Tidy data

```{r}
head(tidy_weather)
```

#### Cleaning and preparing the data

First thing I'm going to delete the "X" before the number of days in day column and concatenate day, month and year column into on column for date. I need to change the format to date so it will not be a string anymore.

```{r}
cleaned_weather <- tidy_weather
cleaned_weather[['day']] <- str_replace(cleaned_weather[['day']], "X", "")
cleaned_weather <- unite(cleaned_weather, date, year, month, day, sep = "/")
cleaned_weather[['date']] <- as.Date(cleaned_weather[['date']], "%Y/%m/%d")
head(cleaned_weather)
```

Also, I saw that there are "T" in precipitation column. It means "trace" and we can change this by 0.

```{r}
cleaned_weather[['PrecipitationIn']] <- str_replace(cleaned_weather[['PrecipitationIn']], "T", "0.00")
head(cleaned_weather)
```

And finally change our values' format to numeric ones so I can use summary function after. To do so, I need to first to switch Events and CloudCover columns.

```{r}
cleaned_weather <- select(cleaned_weather, date, Events, CloudCover:WindDirDegrees)
cleaned_weather[, 3:23] <- sapply(cleaned_weather[, 3:23], as.numeric)
head(cleaned_weather)
```

Now that everything is fine I can check if there are missing values.

```{r}
summary(cleaned_weather)
```

We can see that there are 6 NA values in Max.Gust.SpeedMPH, we know thanks to head that there are Empty values in Events too. We can also see that there is a mistake in Max.Humidity column. Humidity is measured by % so the maximum should be 100 instead of 1000.

```{r}
cleaned_weather[["Events"]][cleaned_weather[["Events"]] == ""] <- "none"
cleaned_weather[["Max.Humidity"]][cleaned_weather[["Max.Humidity"]] == 1000] <- 100
```

As I don't know what to do with NA values in Max.Gust.SpeedMPH I decided to leave it like that.

#### The code book with final cleaned dataset

```{r}
head(cleaned_weather)
summary(cleaned_weather)
boxplot(cleaned_weather[, 3:22])
boxplot(cleaned_weather[,23])
```

# Task 4
### function + dplyr

Instructions are to do the same as for task 1 but by using dplyr functions.